\chapter*{Acknowledgements}

When I started my PhD in 2018 some of the other PhD students were finishing theirs.
This gave me the opportunity to go through their dissertations and get a sneak peek of what was expected of me.
As one could imagine, their work was outstanding and, most impressively, of the highest scientific standard.
Immediately, I thought to myself that I would never be able to achieve something like that.

After four years I can say for certain that I was indeed right, I could never finish this journey -- not alone!
Therefore, I would like to spend the next few paragraphs expressing my most sincere gratitude to all the people who supported me.

First and foremost, I would like to thank my two supervisors, Prof. Paris Avgeriou and Prof. Francesca Arcelli.
Dear Paris, I could not ask for a better supervisor and mentor. Your way of thinking and approaching science has helped me not only become a better researcher but also a better person. 
Your encouragement and guidance helped me overcome obstacles I deemed insurmountable.
I wish I could be half as good of a scientist as you are, but I will be happy even being half as good as you think of me. Thank you.

Dear Francesca, you always believed in me and were always there to help whenever I needed a hand. This holds not only for my PhD but also for my MSc and BSc theses.
You always believed in me and encouraged me to do more and better. I want to deeply thank you for your mentorship and all the opportunities you gave me. I wouldn't have done this without you.

I would like to sincerely thank the assessment committee, Prof. Andrea Capiluppi, Prof. Serge Demeyer, and Prof.  Gregorio Robles, for taking the time to read my thesis and providing me with valuable feedback.

I am also grateful to my brother, Cezar Sas. Cezar, I deeply appreciate that we managed to travel, explore, bike, hike, chat and cook together over the past years. I genuinely treasure these moments and I hope we can share many more in the upcoming years! 

I would also like to thank Jie Tan. Dear Jie, we started our PhD journey together, and, as you would like to jokingly point out, you finished much earlier than me! To be honest, I never stood a chance against you! 
Seeing your commitment and sheer will facing setbacks was inspiring and helped me endure mine.
I am glad we sailed together the sea of adversities entailed by a PhD, and I want you to know that I hold dear to the lifelong memories of our wonderful travels, food, and wine we shared.

Another person I must absolutely thank is Francesco Olivari. Dear Francesco, you are one of the most sincere and genuine people I know.  
I wish we could meet and share a burger -- rigorously at Wereldburgers -- like we used to do during our time in Groningen.
Thank you for hanging out with me! 

Dear Ilaria, I am so grateful we shared the same research interests and got the chance to work on the same research topic as you. I am sure that our future professional endeavours will be more than successful!

Dear Andrea, Yuri, and Ottavio. You are my lifelong friends, and I am so grateful for your support throughout these years. 
After four years spent abroad, I cannot wait to get the chance to hang out with you and watch (and discuss) movies, music, and video games!

Dear Simone, Marco C., Matteo, Marco G., Kristina, Susanna, and Marco De G., thank you for the amazing time we spent together. It is always nice joining you for dinner and other social activities!

Dear Bozhidar, you have no clue how much you helped me throughout this PhD. You cheered me up during my lows and helped me look forward with a positive mindset.

Naturally, I must also express my gratitude to the people at the Software Engineering group (SEARCH) of the Bernoulli Institute that I have not mentioned earlier: Dr. Vasilios Andrikopoulos,  Dr. Ayushi Rastogi, Dr. Daniel Feitosa,  Dr. Mohamed Soliman, H\'ector Cadavid, Ruyin Li, Yikun Li, João Biazotto, Zaki Pauzi, Sara Mahdavi-Hezavehi, and Anja Reuter. 
I am honoured I worked with you, and, every so often, shared a beer with you!

Last but not least, I wish to state my utmost appreciation to my parents.
Their unconditional support and sacrifice are what made me the person I am today.


\begin{flushright}
    Darius Sas\\
    Groningen\\
    August 11, 2022
\end{flushright}


\begin{flushright}
    Darius Sas\\
    Groningen\\
    August 11, 2022
\end{flushright}