\chapter{Supplementary material for Chapter \ref{chap:2}}\label{c2:appendix}
\section{Interview Guide}
\begin{itemize}
    \item Start of the interview [3 min]
    \begin{itemize}
        \item Introduce yourself, your job, and your goal
        \item Briefly mention what this interview is for and disclaimer on how their responses are used and that the interview is confidential
        \item Mention that he is free to expand on any topic or anecdote
        \item Ask permission to record 
        \item Start recording
    \end{itemize}

    \item Background information [3 min]
    \begin{itemize}
        \item What is your current official position? 
        \item What is your role in your team? (day-to-day tasks example) 
        \item How many years of experience do you have in the current position and in total? 
    \end{itemize}

    \item Explanation [3-5 min]
    \begin{itemize}
        \item Explanation of Architectural Smells and negative effects on maintenance activities (concise and stick to literature) 
    \end{itemize}

    \item RQ1 questions [5 min]
    \begin{itemize}
        \item Which of these smells are you already familiar with, from your work? 
        \item How many of them are there in the system you currently work on?
        \item What types of smells do you think are the most important in your case? Why? 
    \end{itemize}

    \item RQ2 questions [5 minutes]
    \begin{itemize}
        \item What types of smells do you deem to be more detrimental for the Maintainability of the system? How are they detrimental?
        \item Can you remember experiencing any issue while maintaining a class/package that could be related to an architectural smell?
        \item In your project are there  obstacles to the implementation of new features that you think are related to architectural smells? If yes, can you tell us about these obstacles?
    \end{itemize}

    \item RQ3 questions [5 minutes]
    \begin{itemize}
        \item Did you try to refactor the part involved in the smell? If yes, how did you do it? If not, why?
        \item Are there any practices in your team to manage the smells and their consequences?
        \item Are you using any tools to monitor architectural issues (or at least dependency analysis)? 
        \item If yes, why did you choose that tool in particular? If not, are you aware of such tools that you would consider using? 
        \item What do you think is the ideal (imaginary) tool that could handle architectural smells issues?
    \end{itemize}
\end{itemize}

\section{Tables}

\begin{table}[h]
    \centering
    \footnotesize
    \caption{Desired features of an ideal tool to manage AS with the reported frequency in parentheses.}
    \label{c2:tab:ideal-features}
    \begin{tabular}{@{}ll@{}}\toprule
    \textbf{Classification} & \textbf{Description} \\ \midrule
    Detection & Highlight the problem while I am writing the code (to save time) (2) \\
     & Automatically detect HL and CD (1) \\
     & Dive deep down the layers and show the actual causes of the smell (2) \\
     & Show the files that create the dependencies of a certain package (1) \\
    Refactoring support & Refactoring suggestions (4) \\
     & What-if analysis to remove cycles (1) \\
     & Show the data or call dependencies between artefacts (1) \\
     & Support to split GC by features (1) \\
    Prioritization & Filter the components to show (2) \\
     & Heatmap of smells in the system (1) \\ \bottomrule
    \end{tabular}
\end{table}