\chapter{Supplementary material for Chapter \ref{chap:4}}
\section{Interview guide}\label{c4:appendix:interview-guide}
\subsection{Questions}
This section lists the questions for the steps listed in the outline.
Total duration of the interview: \textbf{35 minutes} max.

\paragraph{Introduction (3 min.)}
\begin{enumerate}
    \item Introduce yourself, your job, and your goal.
    \item Briefly mention what this interview is for and disclaimer on how their responses are used.
    \item Feel free to expand on any topic or anecdote.
\end{enumerate}

\paragraph{Background information (2 min.)}
\begin{enumerate}[resume]
    \item What is your current official position?
    \item What is your role in this project? (day-to-day tasks example)
    \item How many years of experience do you have in the current position and in total?
\end{enumerate}

\paragraph{Results presentation (8 min.)}
\begin{enumerate}[resume]
    \item Share screen.
    \item Explanation of Architectural Smells and negative effects on maintenance activities (concise and stick to literature)
    \item Explanation in detail of each section of the report by going through each table and plot.
\end{enumerate}

\paragraph{General (2 min.)}
\begin{enumerate}[resume]
    \item General insights emerging from the analysis
    \begin{enumerate}
        \item What are one negative and one positive aspects about the FC that emerged after inspecting the architecture analysis results?
%        \item Did you notice any particular pattern emerging from the results? If so, what information did you use?
    \end{enumerate}
\end{enumerate}

\paragraph{RQ4 (10 min.)}
\begin{enumerate}[resume]
    \item Importance of smell types and characteristics
    \begin{enumerate}
        \item What type of smell do you think is the most important in your case? Why?
        \item For each smell type we calculate specific metrics (called smell characteristics),
        which one, for each smell type, is of most interest for you? Why?
    \end{enumerate}
    \item Perceived and actual quality of the system
    \begin{enumerate}
        \item Did your perception of the quality of the components change after inspecting the results? How/Why not? If yes, can you make an example?
        \item Does the presence of smells confirm what you already knew?
        \item Do the smells affect parts of the system that you were expecting to have issues with? Why/How come?
        \item Are there any missing parts in our analysis?
    \end{enumerate}
\end{enumerate}

\paragraph{RQ5 (10 min.)}
\begin{enumerate}[resume]
    \item Impact of smells on Maintainability
    \begin{enumerate}
        \item What types of smell do you deem to be more detrimental for the Maintainability of the system?
        \item Can you give an example of an issue you experienced while maintaining a component affected by a smell?
        \item Do you think it was related with the presence of a smell?
        \item Would it be hard to fix these issues? What aspects of the smell make it hard to do so (size, affected elements, overlaps)?
    \end{enumerate}
    \item Impact of smells on Evolvability (e.g. how easy it is to implement new features)
    \begin{enumerate}
        \item How have these issues affected the Evolvability (implementation of new functionality) of the affected parts?
        \item (If not answered before) Do you remember any issue hindering the addition of new features to any of the components affected by smells?
        \item Have you been discussing specific obstacles for the implementation of new features?
    \end{enumerate}
    \item Possible remediation strategies
    \begin{enumerate}
        \item What would be a possible quality-improvement plan that you could implement based on the information acquired from this report? And what would help you implement it?
        \item Do the results help you prioritising the issues to fix? If yes, how/why? If no, what could help?
    \end{enumerate}
\end{enumerate}
\paragraph{Feedback (2 min.)}
\begin{enumerate}
    \item General Feedback on the results
    \begin{enumerate}
        \item What view provided you with the most valuable insights (Smell characteristics, Dependency graph, DSM)? Why?
    \end{enumerate}
\end{enumerate}
    