\documentclass[b5paper]{article}
\usepackage[margin=0.7in,top=-.3in,bottom=0in]{geometry}
\usepackage{palatino}
\usepackage{enumerate}
\usepackage{xcolor}

\author{\Large{Darius Sas}} % TODO Update Title
\title{\Large{\bfseries Stellingen}\\\large{behorende bij het proefschrift}\\[.2cm]
\Large{\emph{Managing Technical Debt:\\prioritising and quantifying architectural smells}} \\[.2cm]{\large{van}}}
\date{}

\begin{document}
\maketitle
\thispagestyle{empty}

\begin{enumerate}
\setlength{\itemsep}{3mm}
\begin{large}

\item The architectural smells detected in open source systems evolve rapidly: 50\% of Cyclic Dependencies are replaced by a new instance within the next 3 releases. {\color{teal}{Chapter 2}}

\item The number of artefacts affected by a single Cyclic Dependency instance tends to stay constant over time. The opposite holds for Hublike Dependency smell, which grows in 65\% of the cases. {\color{teal}{Chapter 2}}

\item Software practitioners rank change propagation as the most disruptive effect on maintainability associated with architectural smells. Difficulty in understanding and properly orient in the code follow next. {\color{teal}{Chapter 3}}

\item The God Component and Hublike Dependency smells were considered the most severe smells. {\color{teal}{Chapter 3}}

\item As opposed to open source systems, architectural smells in industrial systems of all types exhibit longer lifespans. {\color{teal}{Chapter 4}}

\item In industrial software systems, Cyclic dependencies are precursors to other, more severe architectural smells. Conversely, Hublike dependencies are the most common successor. {\color{teal}{Chapter 4}}

\item Source code artefacts affected by architectural smells are subject to  more frequent and larger changes than non-affected artefacts.  {\color{teal}{Chapter 5}}

\item The introduction of an architectural smell follows with an increase in the change frequency in the affected artefact. {\color{teal}{Chapter 5}}

\item Under certain conditions, applying large refactorings in long-lived systems is almost impossible for practitioners, as they would risk compromising vital run-time quality attributes. This limits the use of all TD management tools.  {\color{teal}{Chapter 6}}

\item Trade-offs between design-time and run-time quality attributes are very common, but they are often implicit, due to the lack of adequate monitoring tools and practices. {\color{teal}{Chapter 7}}


\end{large}

\end{enumerate}

\end{document}
